How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-07-20 3.08-References.Rmd
Add files via upload
# After copy edits

Having the references at the end of your chapter or publication is actually very important because in-text citations are not sufficient to determine exactly which paper you are referring to. Once you become very familiar with your own specific field you will find that you do glance at the references to make sure that the authors are citing the papers you think they are. It's also a great place to learn about literature that you don't know about already, and for this you need the full citation in order to look up the paper.

The right time to make sure that references are in the document is when you are writing the document. Do not leave this until the end. It has happened to me many times that I have written an in-text citation, did not paste the citation of the end of the document, and by the time I had finished writing, had forgotten what I had read and where I had read it. This then leads to a lot more time wasting trying to find that or another relevant citation.

[Earlier in this book](citations.html), I provided a guide to the differences between Vancouver and Harvard referencing styles. Here I provide the output for the references that were given in that section to demonstrate the way in which these different referencing styles take up very different amounts of space. In that section, I provided some quotes from a published paper that had 9 citations. Below, you can see what these 9 citations look like in the two different generic outputs: Vancouver and Harvard.

## Vancouver style {#vancouver}
The following references follow the text written in [Vancouver style](citations.html) in the section on citations. Note that the order of citations here corresponds to the order in which they are cited.

>^1^Wake DB, Vredenburg VT. Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_. 2008;105(Supplement 1):11466–73.
>
>^2^Collins JP, Crump ML, Lovejoy III TE. _Extinction in our times: global amphibian decline_. Oxford University Press; 2009.
>
>^3^Pimm SL, Jenkins CN, Abell R, Brooks TM, Gittleman JL, Joppa LN, et al. The biodiversity of species and their rates of extinction, distribution, and protection. _Science_. 2014;344(6187):1246752.
>
>^4^Kupferberg SJ. Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_. 1997;78(6):1736–51.
>
>^5^Dufresnes C, Dubey S, Ghali K, Canestrelli D, Perrin N. Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_. 2015;16(6):1507–13.
>
>^6^Berger L, Speare R, Hyatt A. _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_ Environment Australia, Canberra. 1999;1999:23–33.
>
>^7^Daszak P, Cunningham AA, Hyatt AD. Infectious disease and amphibian population declines. _Diversity and Distributions_. 2003;9(2):141–50.
>
>^8^La Marca E, Lips KR, Lotters S, Puschendorf R, Ibanez R, Rueda-Almonacid JV, et al. Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_. 2005;37(2):190–201.
>
>^9^Martel A, Spitzen-van der Sluijs A, Blooi M, Bert W, Ducatelle R, Fisher MC, et al. *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_. 2013;110(38):15325–9.
>
>

## Harvard style {#harvard}
The references below are exactly the same as those above, but follow the [Harvard](citations.html) referencing format. The order of these references is alphabetical. If a paper has the same authors the oldest one usually comes before subsequent papers.

@@

Be aware that these styles (Vancouver and Harvard) represent two broad approaches, every publisher has their own variation on these, and that they can vary substantially between different journals. Full details are always provided by the journal (in their **Instructions to Authors**), and you can refer to these to see how your references need to be formatted. I also provide some pointers towards using [reference managers](referencesoftware.html). These can save you a lot of time, if you've already invested time in setting them up. Alternatively, they can be very frustrating.

@@

There are lots of better things that you could be doing with your time instead of formatting references. And so I'm not going to spend a long time here telling you that you must do it right. I really believe that on first submission of your manuscript you should be able to submit your references in any format. it really only makes sense for you to do all the work formatting them if you aren't going to pay to have published the paper. An increasing number of journals do allow you to use any referencing style (within the generic Vancouver or Harvard format) when you first submit. You are only then required to format them if your ms is likely to be accepted.

There are lots of anal academics out there who will delight in looking for every missing full-stop, comma and capital-letter missing from your references. Perhaps it's because they feel secure pointing out these errors, but are incapable of knowing whether or not you've written good or bad science. Please don't join them and if you can look for a better world where you spend your valuable time doing more important things.

@@

However, DOIs cannot replace references, otherwise we'd need to be able to click on every link all the time, and couldn't read any paper without a connection to the internet. It's still really useful to be able to read a formatted reference at the end of a paper.

### DOI Tools

>>>>

Having the references at the end of your chapter or publication is actually very important because in-text citations often are not sufficient to determine exactly which paper you are referring to. Once you become very familiar with your own specific field you will find that you do glance at the references to make sure that the authors are citing the papers you think they are. It's also a great place to learn about literature that you don't know about already, and for this, you need the full citation in order to look up the paper.

The right time to make sure that references are in the document is when you are writing the document. Do not leave this until the end. It has happened to me many times that I have written an in-text citation, did not paste the citation at the end of the document, and by the time I had finished writing, had forgotten what I had read and where I had read it. This then leads to a lot more time wasting trying to find that or another relevant citation.

[Earlier in this book](#citations), I provided a guide to the differences between Vancouver and Harvard referencing styles. Here I provide the output for the references that were given in that section to demonstrate the way in which these different referencing styles take up very different amounts of space. In that section, I provided some quotes from a published paper that had 9 citations. Below, you can see what these 9 citations look like in the two different generic outputs: Vancouver and Harvard.

## Vancouver style {#vancouver}
The following references follow the text written in [Vancouver style](#citations) in the section on citations. Note that the order of citations here corresponds to the order in which they are cited.

>^1^Wake DB, Vredenburg VT. (2008) Are we in the midst of the sixth mass extinction? A view from the world of amphibians. _Proceedings of the National Academy of Sciences_. 105(Supplement 1):11466–73.
>
>^2^Collins JP, Crump ML, Lovejoy III TE. (2009) _Extinction in our times: global amphibian decline_. Oxford University Press.
>
>^3^Pimm SL, Jenkins CN, Abell R, Brooks TM, Gittleman JL, Joppa LN, et al. (2014) The biodiversity of species and their rates of extinction, distribution, and protection. _Science_. 344(6187):1246752.
>
>^4^Kupferberg SJ. (1997) Bullfrog (*Rana catesbeiana*) invasion of a California river: the role of larval competition. _Ecology_. 78(6):1736–51.
>
>^5^Dufresnes C, Dubey S, Ghali K, Canestrelli D, Perrin N. (2015) Introgressive hybridization of threatened European tree frogs (*Hyla arborea*) by introduced _H. intermedia_ in Western Switzerland. _Conservation Genetics_. 16(6):1507–13.
>
>^6^Berger L, Speare R, Hyatt A. (1999) _Chytrid fungi and amphibian declines: overview, implications and future directions. Declines and disappearances of Australian frogs_ Environment Australia, Canberra. 1999:23–33.
>
>^7^Daszak P, Cunningham AA, Hyatt AD. (2003) Infectious disease and amphibian population declines. _Diversity and Distributions_. 9(2):141–50.
>
>^8^La Marca E, Lips KR, Lötters S, Puschendorf R, Ibanez R, Rueda-Almonacid JV, et al. (2005) Catastrophic population declines and extinctions in Neotropical harlequin frogs (Bufonidae: *Atelopus*) _Biotropica: The Journal of Biology and Conservation_. 37(2):190–201.
>
>^9^Martel A, Spitzen-van der Sluijs A, Blooi M, Bert W, Ducatelle R, Fisher MC, et al. (2013) *Batrachochytrium salamandrivorans* sp. nov. causes lethal chytridiomycosis in amphibians. _Proceedings of the National Academy of Sciences_. 110(38):15325–9.
>
>

## Harvard style {#harvard}
The references below are exactly the same as those above, but follow the [Harvard](citations.html) referencing format. The order of these references is alphabetical. If a paper has the same authors the oldest paper usually comes before subsequent papers.

@@

Be aware that these styles (Vancouver and Harvard) represent two broad approaches, every publisher has their own variation on these, and that they can vary substantially between different journals. Full details are always provided by the journal (in their **Instructions to Authors**), and you can refer to these to see how your references need to be formatted. I also provide some pointers towards using [reference managers](#referencesoftware). These can save you a lot of time, if you've already invested time in setting them up. Alternatively, they can be very frustrating.

@@

There are lots of better things that you could be doing with your time instead of formatting references. And so I'm not going to spend a long time here telling you that you must do it right. I really believe that on first submission of your manuscript you should be able to submit your references in any format. it really only makes sense for you to do all the work formatting them if you aren't going to pay to publish the paper. An increasing number of journals do allow you to use any referencing style (within the generic Vancouver or Harvard format) when you first submit. You are only then required to format them if your ms is likely to be accepted.

There are lots of anal academics out there who will delight in looking for every missing full-stop, comma and capital letter missing from your references. Perhaps it's because they feel secure pointing out these errors, but are incapable of knowing whether or not you've written good or bad science. Please don't join them and, if you can, look for a better world where you spend your valuable time doing more important things.

@@

However, DOIs cannot replace references, otherwise, we'd need to be able to click on every link all the time, and couldn't read any paper without a connection to the internet. It's still really useful to be able to read a formatted reference at the end of a paper.

### DOI tools

### 2021-07-20 3.09-Tables.Rmd
Add files via upload
# After copy edits

## Tables {#tables}

@@

| **Do** | **Don't** |
|------|------|
| Use a table as a means of reducing the length of your text. |  Merge cells, especially in irregular ways, as this makes formatting and copy/paste of your table prone to errors. |
| Make your headings clear and concise. | Omit units of measurement |
| Summarise extensive data. | Try to include every data point (that's what [Supp Info](SuppInfo.html) is for). |
| Use tables to provide repeated statistical information on tests or models. | Feel the need to include all of these in the ms  (that's what [Supp Info](SuppInfo.html) is for). |
| Provide a full table legend that explains exactly what is in it (see [figures](figures.html)). | Make tables so large that they have to run over several pages. If this can go into [Supp Info](SuppInfo.html) then it should. |
| Be adventurous by using tables to produces ideas, even with entire paragraphs of text or photographs. | Feel obliged to stick to using tables only for numbers. |
| Use shading for every other row. | Have rows or columns so close together that you can't tell them apart. |

>>>>

# Tables {#tables}

@@

| **Do** | **Don't** |
|------|------|
| Use a table as a means of reducing the length of your text. |  Merge cells, especially in irregular ways, as this makes formatting and copy/paste of your table prone to errors. |
| Make your headings clear and concise. | Omit units of measurement |
| Summarise extensive data. | Try to include every data point (that's what [Supp Info](#suppInfo) is for). |
| Use tables to provide repeated statistical information on tests or models. | Feel the need to include all of these in the ms  (that's what [Supp Info](#suppInfo) is for). |
| Provide a full table legend that explains exactly what is in it (see [figures](#figures)). | Make tables so large that they have to run over several pages. If this can go into [Supp Info](#suppInfo) then it should. |
| Be adventurous by using tables to produces ideas, even with entire paragraphs of text or photographs. | Feel obliged to stick to using tables only for numbers. |
| Use shading for every other row. | Have rows or columns so close together that you can't tell them apart. |

### 2021-07-20 3.10-Figures.Rmd
Add files via upload
# After copy edits

## Figures

I consider myself to be a visual person, somebody who understands more by looking at pictures than by reading words. Not everyone is the same but there are a lot of people who can derive more information from a good figure. Therefore for me, a figure is an opportunity to convey maximum amounts of information. It can be a relatively simple figure like a graph or bar chart, a map, a conceptual diagram, or my favourite, a composite figure.

The way in which to visualise your data requires another book on the topic, as there are a myriad of decisions to make on all sorts of topics. What I provide below is a very brief start to thinking about these topics. If you need to explore more of the world of data visualisation, I suggest that you read books on this topic. Happily, people like Kieran Healy [-@healy2018data], Jack Dougherty and Ilya Ilyankou [-@dougherty2021hands-on] and Claus Wilke [-@wilke2019fundamentals] have each written such books. Claus points out that while there are ways in which you can go wrong when plotting data (such as drawing lines to infer relationships that are not there), the skills in data visualisation are more in the aesthetic and interpretive environment of what tools to use that make the data easiest to interpret. How will the figure enhance understanding of the data?

Note that the principal software used for drawing data in Dougherty & Ilyankou's, Kieran’s and Claus’ books is R ([see part 1](useR.html)), and the many packages that have been developed for use within the R environment [@rcoreteam2021r]. The distinct advantages that this produces are the almost limitless additions and details that you can make to these figures to produce exactly what you want, in a [vector format](vector.html). The drawback is that you might need to invest some time to learn how the graphics package works, and how to manipulate it to do what you want.

@@

### Graphs

You should be aware of all the things that make up a good graph already. here I'm going to provide some extra pointers on what I think is important and should be considered when drawing a good graph:

- __Labels__. This seems completely obvious and yet it is a common problem with submitted papers. Axes should be labelled with both an explanation and the units in which they were measured. 
  + __Capitals__. Be consistent in capitalising labels and other words in your figure. Avoid using just capitals (e.g. ALL CAPITALS ARE DIFFICULT TO READ AND UNSIGHTLY)
- __The scale__. make sure that the numbers of the scale are large enough to read even when the figure is reduced the size that it will be printed on the page. This includes the text that crossed the axes.
  + If you have used a log scale, make sure that this is prominently labelled.
- __Data points__. I am a great fan of including all the data points that go to make up any summary statistic that is displayed on the graph. Thus, if you provide a mean or something like a box plot for your x-axis category, then please add all data points as well. This can be done using the jitter function in R. 
- Don’t draw relationships that aren’t there. The ease of plotting graphs makes it possible to display relationships between data points that don’t exist. This is most commonly done across a set of discrete measurements when a continuous line links them without any data to suggest this relationship. Be especially careful when plotting lines that you can infer a continuous value between points plotted.
- Compiling graphs. if you have a series of graphs that share the same x or y axis then consider compiling them together. This essentially turns them into a composite figure, but it often means that readers can compare data easily from one graph to another. remember that you will have to make sure that the scales on the shared access are all the same. 
- Use consistent symbols and colours throughout your thesis. Your thesis will likely plot data on similar variables in many ways throughout. By using consistent symbols (e.g. circles for females and triangles for males) and colours throughout the thesis for the variables that chapters have in common, you will help your reader understand a figure much more quickly. The consistency of the graphical environment of a thesis will facilitate understanding.
  + Consistency is especially important across panels within a composite figure.
- Colour. The reason for all the fuss about not having colour in figures is mostly to do with the extra cross cost involved in printing them. Typically these days it is possible to use colour in figures and for these to be reproduced without any extra charge online. Thus if the data that you have is best shown with colour then use colour. There are some things to be mindful of when using colour.
  + Colour blindness: This affects quite a large portion of the population, particularly men. To avoid using colours that can be easily confused simply look them up for example here
  + Printing colour in monochrome: some people still like to print out papers in order to read them, annotate them or share them at a journal club. using a monochrome printer many colours are indistinguishable. you can therefore change the hue or intensity of a colour to make sure that it can be distinguished when it is printed. 
  + Aesthetic quality: some colours don't go well together. a lot of people have already determined this and you can find a good guide [here](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3).

### Maps
There are many reasons why you might need to include a map in your paper when writing for biological sciences. I find that the easiest way of creating Maps is to use GIS software like QGIS or ArcGIS. Both of these require a good degree of learning in order to master. To have a map with sampling points is relatively straight forward. Especially with QGIS you can watch ‘how to’ videos on the internet to learn most of what you need to know. things to remember when you're drawing a map:
If your map is a detailed view of an area then consider a small context map in the corner showing where your area is
- Include the scale
- Provide a North arrow especially if you have rotated the view
- Some journals require a specific mention of the geographic projection [see here](https://www.gislounge.com/map-projection/), or to provide a citation for each layer of the map (which makes sense when you think about it). Thus it’s a good policy to make sure you know the provenance of the layers that you are using.
- When using colour see [these](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) handy sets of colours to use at Colorbrewer.
- Make sure that any points you have included are big enough so that you can see the point and the shape of a point when the figure is reduced
- Strictly speaking maps in figures should not have embedded legends (keys). The correct way to explain symbols is to do it in the figure legend. However, sometimes if there are many different symbols it is more expedient to use an embedded legend.

#### Maximising the content of your figure
Sometimes you will have the option of including more information in a figure than the original plan that you started with. This is perhaps especially true when it comes to producing maps with GIS, because you have the option of adding extra data layers. But remember that this information should not be gratuitous. In figure \@ref(fig:not-fancy), we can see a map that shows a number of sampling sites. The map has a scale bar and north arrow as needed.

@@

However, it is relatively simple to add more information to this map in the form of layers. In figure \@ref(fig:fancy-map), a digital elevation model (DEM) allows us to see the relative elevation at the sampling points, while transparent colours provide information about the different biomes in which the samples were made. Lastly, I've used a piece of sea to insert an image of the species sampled. There is also a reference map (top left) showing the area of southern Africa where the sampling points occur. This figure is now much more informative, and has greater aesthetic appeal to me, but all of this added information is only of use if relevant to the manuscript or chapter.

(ref:fancy-map) **Maps tend to have a lot of space, and GIS provides the opportunity to add layers to give rich information, as well as improving aesthetics.** This map shows lots more information about the biomes in which each of the sample sites (red dots) falls, the elevation and an image of the species sampled. Inset of Africa (top left) shows the area sampled with a red square. I've also managed to fit in an image of the species sampled into the sea (bottom right).

@@

### Conceptual diagram
Conceptual diagrams or figures are very useful in the introduction or methods when you have undertaken a complex study design. There are many different ways of doing this, but essentially the information needs to be conveyed in a way that is more simple in the figure than writing it out. Boxes and arrows are sufficient for most purposes. If the figure gets too complicated to understand, it's probably of little use.

I encourage my students to draw conceptual diagrams of how the thesis chapters relate to one another, and then to include this diagram in the introduction to the thesis to help examiners and any other readers to understand how the chapters are interrelated.

@@

- Keep any text in the diagram large enough so that when it is reduced to printing size it can still be read
- Try not to write full sentences but just notes
- Make use of arrows to show how ideas are related
- Look at other conceptual diagrams especially those that resonate with you as a reader

### The composite figure
For me, composite figures started when I noticed other papers were including small pictures of species in phylogenies. As a reader I really liked this way of understanding more about the animals that were being studied. The composite figure also comes about by making use of blank spaces that naturally occur within figures. Many graphs have large white spaces toward the top (left or right), and small images of study animals, cartoons of study designs or systems can be inserted to improve the immediate recognition of what the figure is portraying.

#### Ideas for composite figures:
include a picture of the subject species in a small corner of blank space in one of your figures. For example, maps with a large area of sea (If your subject is a terrestrial species) are especially good for this (see Figure \@ref(fig:fancy-map)).

### Sending your graphics file to the publisher
One consistent problem is the lack of understanding that authors have for different file types. I can be rather confusing as these are stated in lots of different ways, and often in units that are not intuitive. Journals usually have long and detailed explanations of how to do this in their instructions to authors, yet this is one of the most common difficulties that authors have. I will provide a brief overview of the most common problems, but better overviews exist and I encourage you to read them [e.g. @wilke2019fundamentals].

@@

Most publishers will require you to name the files of your figures to correspond with their instructions to authors. If they don't have specific requirements, then use clear file names (e.g. Figure_1.png), and not cryptic ones. Remember that each figure needs a single file, so if you have produced a composite figure, then you will need to output this as a single graphics file.

#### Vectors vs bitmaps {#vector}
The biggest difference among image types are the difference between vectors (a line drawn between two points) and bitmaps (using pixels to create images). Vector diagrams will produce superior images on screen that will allow you to zoom in without loss of resolution (e.g. svg, eps, pdf). With a bitmap, the more you zoom, the more you will be able to see the pixels that make up the figure. If at all possible, create your figures using vector graphics (such as in R or drawing software such as Adobe Illustrator), and store them in this format. The only exceptions will be images (e.g. png, jpg, tif, raw, gif, bmp) that you add to composite figures, insert these with the highest resolution available.

@@

#### Sizing your figure
When you export your graphic from the graphics package that you are using to draw it, specify the resolution (most often in dpi: dots per inch) given by the publisher. Minimum resolution for photos and tone images is 400 dpi, and 600 dpi for line drawings. Remember that the dpi will change if the size of your figure changes. Hence you must know the printed size of your figure (in cm or inches) on the page. This is usually either one or two columns. Rarely, a journal might allow you to use landscape instead of portrait for a figure. As a rule, you should use the page setup in your package to specify the output size of your figure (page layout function) before you start, and then set the resolution before you export the figure. It’s always possible to draw at a higher resolution and convert to a lower resolution. Low resolution images need to be redrawn to make them higher resolution.

@@

### The figure legend
The most important point with the figure Legend is that it should be a stand-alone explanation of what is in the figure. The reader should not have to refer to the text in order to understand what is in the figure. For example, if the figure is about a species then species name should be included in full. The explanation of any variables or categories needs to be provided. But it should not be taken to absurd levels, is not a reason to repeat everything that's in the materials and methods. Take a look at some figure legends that are already published in the journal that you are hoping to submit to and you should have a good idea of what to write. Figure legends need to be full, grammatically correct sentences.

Remember that the figure legend is placed underneath the embedded figure in your thesis. This means that the itself graphic should not have any title. Traditional submissions to journals require figure legends to be at the end of the manuscript, before the figures themselves. Many journals now allow you to embed a lower resolution image of the figure on first submission together with the legend underneath in order to facilitate peer review. Check on the requirements for you institution or journal before finalising this.

A trend with online journals is that each figure (and table) needs a topic or title sentence, followed by a description. This is a good way forward with any figure legend. Both need to be uploaded along with the image file when you submit the manuscript.

>>>>

# Figures

The way in which to visualise your data requires another book on the topic, as there are a myriad of decisions to make on all sorts of topics. What I provide below is a very brief start to thinking about these topics. If you need to explore more of the world of data visualisation, I suggest that you read books on this topic. Happily, people like Kieran Healy [-@healy2018data], Jack Dougherty and Ilya Ilyankou [-@dougherty2021hands-on] and Claus Wilke [-@wilke2019fundamentals] have each written such books. Claus points out that while there are ways in which you can go wrong when plotting data (such as drawing lines to infer relationships that are not there), the skills in data visualisation are more in the aesthetic and interpretive environment of what tools to use that make the data easiest to interpret. How will the figure enhance understanding of the data and respond to your [hypotheses](#hypothesis)?

Note that the principal software used for drawing data in Dougherty & Ilyankou's, Kieran’s and Claus’s books is R ([see part 1](#useR)), and the many packages that have been developed for use within the R environment [@rcoreteam2021r]. The distinct advantages that this produces are the almost limitless additions and details that you can make to these figures to produce exactly what you want, in a [vector format](#vector). The drawback is that you might need to invest some time to learn how the graphics package works, and how to manipulate it to do what you want.

@@

## Graphs

You should be aware of all the things that make up a good graph already. Here I'm going to provide some extra pointers on what I think is important and should be considered when drawing a good graph:

- __Labels__. This seems completely obvious and yet it is a common problem with submitted papers. Axes should be labelled with both an explanation and the units in which they were measured. 
  + __Capitals__. Be consistent in capitalising labels and other words in your figure. Avoid using just capitals (e.g. ALL CAPITALS ARE DIFFICULT TO READ AND UNSIGHTLY)
- __The scale__. Make sure that the numbers of the scale are large enough to read even when the figure is reduced the size that it will be printed on the page. This includes the text that crossed the axes.
  + If you have used a log scale, make sure that this is prominently labelled.
- __Data points__. I am a great fan of including all the data points that go to make up any summary statistic that is displayed on the graph. Thus, if you provide a mean or something like a box plot for your x-axis category, then please add all data points as well. This can be done using the jitter function in R. 
- __Don’t draw relationships that aren’t there__. The ease of plotting graphs makes it possible to display relationships between data points that don’t exist. This is most commonly done across a set of discrete measurements when a continuous line links them without any data to suggest this relationship. Be especially careful when plotting lines that you can infer a continuous value between points plotted.
- __Compiling graphs__. If you have a series of graphs that share the same x- or y-axis then consider compiling them together. This essentially turns them into a composite figure, but it often means that readers can compare data easily from one graph to another. Remember that you will have to make sure that the scales on the shared access are all the same. 
- __Use consistent symbols and colours throughout your thesis__. Your thesis will likely plot data on similar variables in many ways throughout. By using consistent symbols (e.g. circles for females and triangles for males) and colours throughout the thesis for the variables that chapters have in common, you will help your reader understand a figure much more quickly. The consistency of the graphical environment of a thesis will facilitate understanding.
  + Consistency is especially important across panels within a composite figure.
- Colour. The reason for all the fuss about not having colour in figures is mostly to do with the extra cost involved in printing them. Typically these days, it is possible to use colour in figures and for these to be reproduced without any extra charge online. Thus if the data that you have is best shown with colour then use colour. There are some things to be mindful of when using colour.
  + Colour blindness: This affects quite a large portion of the population, particularly men. To avoid using colours that can be easily confused simply look them up.
  + Printing colour in monochrome: some people still like to print out papers in order to read them, annotate them or share them at a journal club. Using a monochrome printer many colours are indistinguishable. You can therefore change the hue or intensity of a colour to make sure that it can be distinguished when it is printed. 
  + Aesthetic quality: some colours don't go well together. A lot of people have already determined this and you can find a good guide [here](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3).

## Maps
There are many reasons why you might need to include a map in your paper when writing for biological sciences. I find that the easiest way of creating Maps is to use GIS software like QGIS or ArcGIS. Both of these require a good degree of learning in order to master. To have a map with sampling points is relatively straight forward. Especially with QGIS you can watch ‘how to’ videos on the internet to learn most of what you need to know. Things to remember when you're drawing a map:

- If your map is a detailed view of an area then consider a small context map in the corner showing where your area is
- Include the scale
- Provide a North arrow especially if you have rotated the view
- Some journals require a specific mention of the geographic projection [see here](https://www.gislounge.com/map-projection/), or to provide a citation for each layer of the map (which makes sense when you think about it). Thus it’s a good policy to make sure you know the provenance of the layers that you are using.
- When using colour see [these](https://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) handy sets of colours to use at Colorbrewer.
- Make sure that any points you have included are big enough so that you can see the point and the shape of a point when the figure is reduced
- Strictly speaking maps in figures should not have embedded legends (keys). The correct way to explain symbols is to do it in the figure legend. However, sometimes if there are many different symbols it is more expedient to use an embedded legend.

### Maximising the content of your figure
Sometimes you will have the option of including more information in a figure than the original plan that you started with. This is perhaps especially true when it comes to producing maps with GIS because you have the option of adding extra data layers. But remember that this information should not be gratuitous. In Figure \@ref(fig:not-fancy), we can see a map that shows a number of sampling sites. The map has a scale bar and north arrow as needed.

@@

However, it is relatively simple to add more information to this map in the form of layers. In Figure \@ref(fig:fancy-map), a digital elevation model (DEM) allows us to see the relative elevation at the sampling points, while transparent colours provide information about the different biomes in which the samples were made. Lastly, I've used a piece of sea to insert an image of the species sampled. There is also a reference map (top left) showing the area of southern Africa where the sampling points occur. This figure is now much more informative, and has greater aesthetic appeal to me, but all of this added information is only of use if relevant to the manuscript or chapter.

(ref:fancy-map) **Maps tend to have a lot of space, and GIS provides the opportunity to add layers to give rich information, as well as improving aesthetics.** This map shows lots more information about the biomes in which each of the sample sites (red dots) falls, the elevation and an image of the species sampled. The inset of Africa (top left) shows the area sampled with a red square. I've also managed to fit in an image of the species sampled into the sea (bottom right).

@@

## Conceptual diagram
Conceptual diagrams or figures are very useful in the introduction or methods when you have undertaken a complex study design. There are many different ways of doing this, but essentially the information needs to be conveyed in a way that is more simple in the figure than writing it out. Boxes and arrows are sufficient for most purposes. If the figure gets too complicated to understand, it's probably of little use.

I encourage my students to draw conceptual diagrams of how the thesis chapters relate to one another (Figure \@ref(fig:thesis-construction)), and then to include this diagram in the introduction to the thesis to help examiners and any other readers to understand how the chapters are interrelated (see [Part IV](PhDintro.html)).

@@

- Keep any text in the diagram large enough so that when it is reduced to printing size it can still be read
- Try not to write full sentences but just notes
- Make use of arrows to show how ideas are related
- Look at other conceptual diagrams, especially those that resonate with you as a reader

## The composite figure
For me, composite figures started when I noticed other papers were including small pictures of species in phylogenies. As a reader, I really liked this way of understanding more about the animals that were being studied. The composite figure also comes about by making use of blank spaces that naturally occur within figures. Many graphs have large white spaces toward the top (left or right), and small images of study animals, cartoons of study designs or systems can be inserted to improve the immediate recognition of what the figure is portraying.

### Ideas for composite figures:
Include a picture of the subject species in a small corner of blank space in one of your figures. For example, maps with a large area of sea (if your subject is a terrestrial species) are especially good for this (see Figure \@ref(fig:fancy-map)).

## Sending your graphics file to the publisher
One consistent problem is the lack of understanding that authors have for different file types. It can be rather confusing as these are stated in lots of different ways, and often in units that are not intuitive. Journals usually have long and detailed explanations of how to do this in their instructions to authors, yet this is one of the most common difficulties that authors have. I will provide a brief overview of the most common problems, but better overviews exist and I encourage you to read them [e.g. @wilke2019fundamentals].

@@

Most publishers will require you to name the files of your figures to correspond with their instructions to authors. If they don't have specific requirements, then use clear file names (e.g. Figure_1.png), and not cryptic ones. Remember that each figure needs a single file, so if you have produced a composite figure, then you will need to output this as a single graphics file. You will need to send them a generic file type (e.g. eps, jpg) and not something specific to the software that you used (e.g. ai, cdr).

**Colour** is usually free to produce online, but you may need to pay for printed production. Make sure that your figure makes sense in monochrome if you intend the colour to be online only.

### Vectors vs bitmaps {#vector}
The biggest difference among image types is the difference between vectors (a line drawn between two points) and bitmaps (using pixels to create images). Vector diagrams will produce superior images on screen that will allow you to zoom in without loss of resolution (e.g. svg, eps, pdf). With a bitmap, the more you zoom, the more you will be able to see the pixels that make up the figure. If at all possible, create your figures using vector graphics (such as in R or drawing software such as Adobe Illustrator), and store them in this format. The only exceptions will be images (e.g. png, jpg, tif, raw, gif, bmp) that you add to composite figures, insert these with the highest resolution available.

@@

### Sizing your figure
When you export your graphic from the graphics package that you are using to draw it, specify the resolution (most often in dpi: dots per inch) given by the publisher. Minimum resolution for photos and tone images is 400 dpi, and 600 dpi for line drawings. Remember that the dpi will change if the size of your figure changes. Hence you must know the printed size of your figure (in cm or inches) on the page. This is usually either one or two columns. Rarely, a journal might allow you to use landscape instead of portrait for a figure. As a rule, you should use the page setup in your package to specify the output size of your figure (page layout function) before you start, and then set the resolution before you export the figure. It’s always possible to draw at a higher resolution and convert to a lower resolution. Low-resolution images need to be redrawn to make them higher resolution.

@@

## The figure legend
The most important point with the figure legend is that it should be a stand-alone explanation of what is in the figure. The reader should not have to refer to the text in order to understand what is in the figure. For example, if the figure is about a species then species name should be included in full. The explanation of any variables or categories needs to be provided. But it should not be taken to absurd levels, and is not a reason to repeat everything that's in the materials and methods. Take a look at some figure legends that are already published in the journal that you are hoping to submit to and you should have a good idea of what to write. Figure legends need to be full, grammatically correct sentences.

Remember that the figure legend is placed underneath the embedded figure in your thesis. This means that the graphic itself should not have any title. Traditional submissions to journals require figure legends to be at the end of the manuscript, before the figures themselves. Many journals now allow you to embed a lower resolution image of the figure on first submission together with the legend underneath in order to facilitate peer review. Check on the requirements for your institution or journal before finalising this.

A trend with online journals is that each figure (and [table](#table)) needs a topic or title sentence, followed by a description. This is a good way forward with any figure legend. Both need to be uploaded along with the image file when you submit the manuscript.

### 2021-07-20 3.11-Credit.Rmd
Add files via upload
# After copy edits

## Who did what? {#credit}

@@

There are a lot of politics around authorship. How do you decide whether or not someone did enough to be included on an author line. This is a very important point that we cover elsewhere. All that I will say here is that this should have been decided early on in the study. Everyone should know whether or not they are going to be an author at the time that they participate. No one should ever find out but they were not an author by reading their name in the acknowledgements.

### CRediT where it’s due

@@

If you want to know more about each of these roles and how they Are defined in the CRediT taxonomy then please visit the site. You can also ‘claim’ these roles under your [Rescognito profile](https://rescognito.com/), linked to your ORCID account.

Even if the journal that you are submitting to does not use the credit taxonomy I would suggest that you use this as the basis for your saying who did what in the publication. See a blog article on this [here](http://crosstalk.cell.com/blog/giving-authors-the-credit-they-deserve). Otherwise you could end up with what is known as hyperauthorship, a phenomenon evidenced by massive co-authorship levels [@cronin2001hyperauthorship].

### Ordering the author line up 
In Biological Sciences it is traditional for the author who did most of the work to take the first position. For you and your thesis you are most likely to be the first author, certainly for all of your thesis chapters.

The last position is normally for the head of the laboratory in which the work was undertaken. This would usually be your advisor, but could be another collaborator, if for example, you went to do some work in their laboratory. Funding also plays a role here, and if you are unsure about who should be listed last, then ask your advisor.

In between the first and last authors is a bit of a no-man's land. Some people order it by who did most (closest to first), but then there is also a push to be closest to last place (closest to being more senior). Alphabetical also works, but that often benefits or penalises those who have names that regularly appear higher up or lower down in the alphabet. One colleague is always first to suggest alphabetical order as he regularly ends up as last author. If it isn’t part of your thesis, should you still be an author?

>>>>

# Who did what? {#credit}

@@

There are a lot of politics around authorship. How do you decide whether or not someone did enough to be included on an author line. This is a very important point. All that I will say here is that this should have been decided early on in the study. Everyone should know whether or not they are going to be an author at the time that they participate. No one should ever find out but they were not an author by reading their name in the acknowledgements.

## CRediT where it’s due

@@

If you want to know more about each of these roles and how they are defined in the CRediT taxonomy then please visit the site. You can also ‘claim’ these roles under your [Rescognito profile](https://rescognito.com/), linked to your [ORCID](#ORCID) account.

Even if the journal that you are submitting to does not use the credit taxonomy. I would suggest that you use this as the basis for your saying who did what in the publication. See a blog article on this [here](http://crosstalk.cell.com/blog/giving-authors-the-credit-they-deserve). Otherwise, you could end up with what is known as hyperauthorship, a phenomenon evidenced by massive co-authorship levels [@cronin2001hyperauthorship].

## Ordering the author line up 
In Biological Sciences, it is traditional for the author who did most of the work to take the first position. For you and your thesis you are most likely to be the first author, certainly for all of your thesis chapters.

The last position is normally for the head of the laboratory in which the work was undertaken. This would usually be your advisor, but could be another collaborator, if, for example, you went to do some work in their laboratory. Funding also plays a role here, and if you are unsure about who should be listed last, then ask your advisor.

In between the first and last authors is a bit of a no-man's land. Some people order it by who did most (closest to first), but then there is also a push to be closest to last place (closest to being more senior). Alphabetical also works, but that often benefits or penalises those who have names that regularly appear higher up or lower down in the alphabet. One colleague is always first to suggest alphabetical order as he regularly ends up as last author.
